---
title: "Road surface: import and explore"
author: "Mark Ruddy"
date: "2017-02-27"
output: html_notebook
---

Data on road lengths in Great Britain can be found from the [Office for National Statistics (ONS)](https://www.gov.uk/government/statistical-data-sets/rdl02-road-lengths-kms). This is provided as road type by GB Local Authority from 2005 to 2015. A supplement also lists road lengths, with various data gaps, dating back to 1914.

The Ordnance Survey also holds information on road length in their [OS Open Roads](https://www.ordnancesurvey.co.uk/business-and-government/products/os-open-roads.html) product - available under the Open Government Licence ([OGL](http://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/)). This could potentially allow any analysis to be broken down into smaller geographic areas than that achievable from the ONS data. But OS Open Roads only incldues Motorways, A-roads, B-roads.

Both datasets require assumptions on the width's of different road types. The ONS estimate includes OS data and additionally draws on information from Local Authorities, the Scottish and Welsh governments, and the Highways Agency. With the additional information present in the ONS data I decided to go with that as a basis for estimating road area.

## ONS road length

[Office for National Statistics (ONS)](https://www.gov.uk/government/statistical-data-sets/rdl02-road-lengths-kms) are downloadable as .ods files. 

Focussing on the Local Authority dataset:

```{r load-packages, include=FALSE}
library(tidyverse)
library(readODS)
library(httr)
```

I'm interested in looking at road lengths across all years (2005-2015) and 

## The ONS dataset

```{r read-data}

url <- "https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/523715/rdl0202.ods"

GET(url, write_disk("../data/ONS-GB-roads.ods", overwrite = T))

ons.roads <- read_ods("../data/ONS-GB-roads.ods", sheet = 1, skip = 6)
```

Inspecting ONS-GB-roads.ods in OpenOffice shows that:

* There are 6 lines to skip before the dataset begins.
* Sheets contain different types of road length information. Those sheets with names prefixed "RDL0202a"" contain Motorway, 'A' road, and Minor road data broken down by Local authority. These eleven sheets (years) are the data we're interested in.
* The cell range of the dataset (including column names) is:
  * A7:Z211 for 2005 to 2008; 
  * A7:Z213 from 2009 to 2012 
    - Rural 'C' and 'U', and Urban 'C' and 'U' are separate up to 2013.
    - Cheshire UA is split into Cheshire East UA, and Cheshire West and Chester UA from 2009.
    - Bedfordshire is split into Bedford UA and Central Bedfordshire UA from 2009.
  * and A7:X213 from 2013 to 2015 - Rural 'C' and 'U', and Urban 'C' and 'U' are combined from 2013.
* There is a variety of lines of meta-data below cells containing the dataset.



```{r read-r}
path <- "../data/ONS-GB-roads.ods"

## Sheets we're interested in
sheets <- c(seq(1,19,2),20)

ons.call <- function(sheet, range) {
  read_ods(path, sheet = sheet, col_names = F, skip = 6) #skip =6, range = "A7:Z213"
} 

rdl.list <- lapply(sheets, ons.call)
```

Having read data on road length from the .ods sheets we need to clean it up. A number of differences in the dimensions of the dataset were highlighted above. We need to harmonise each list element (sheet) in order to combine them all together and make them 'tidy' for analysis.

I could do all this manipulation in the spreadsheet but the time spent manually tinkering to give the format needed is not time well spent. Some of the operations are repetative and getting them down in R enables them to be applied across all eleven years. Also, it's a good R learning experience.

## Bringing the whole road length dataset together

What are the steps that need to be taken to combine all the road length data together into a single dataset?

1. Add a year column to each list element.
2. 

```{r tidy}
head(rdl.list[[1]])

lapply()


# rdl = do.call("rbind.data.frame", rdl.list) ## combine dfs from list
```












